Geometric variational problems

### 2020-11-12 NSF2020/Description/specificaim1_ver2.tex
RJR

The HK framework assumes a three-dimensional lipid monolayers with internal structure
consisting of straight fibers that represent elongated hydrocarbon chains. HK assumed an
elastic energy density ${\cal W}$ quadratic in the Green-Lagrange strain tensor for membrane deformation ${\bf x}$.
%
%
%
%These fibers
%tilt and stretch with respect to the dividing surface--the surface formed lying between the hydrocarbon chains and polar heads of lipids.
%Accordingly, the deformation takes the form 
%\begin{equation}
%  \label{LMdeformation}
%{\bf x}(X_1, X_2, X_3) = {\bf x}_0(X_1, X_2) + \zeta(x_0(X_1, X_2), X_3) {\bf n}(x_0(X_1, X_2))
%\end{equation}
%where $(X_1,X_2,X_3)$ are points in a reference volume.
%The map $x_0$ parametrizes the dividing surface $\Sigma$, and the unit vector field ${\bf n}$ parametrizes the
%direction of the lipid tails in the deformed state \cite{doi:10.1021/jp075641w,KLAUDA20083074}.
%The parameter $X_3$ and the function $\zeta$ parametrizes the distance along the hydrocarbon chain in the reference state
%and the deformed state, respectively, and so $0 \leq X_3 \leq \delta$ where $\delta = $ 2.5 nm is a realistic monolayer thickness. 
%
%The elastic theory assumes an energy density that is quadratic in the Green-Lagrange strain tensor for ${\bf x}$. 
%The chains stretch and compress to satisfy incompressibility.
%Symmetries with respect to mirror reflection and in-plane rotation, the internal structure assumptions and incompressibility
%lead to an elastic surface energy over the dividing surface $\Sigma$. 
%
This energy decomposes into four, fundamental and independent deformations  (figure~\ref{fig:deformations}):
splay ($\Div {\bf n}$), twist ($\Curl {\bf n}$), saddle splay ($\det \nabla {\bf n}$) and tilt ${\bf t}={\bf n}/({\bf N}\cdot {\bf n}) - {\bf N}$ where ${\bf N}$ is the unit surface normal:
\begin{equation}
\label{ansatz3}
\begin{aligned}
{\cal W} \equiv \int_{\Sigma} 
  \tfrac{1}{2}\KB\left[ \left( \Div {\bf n} + k_0\right)^2 - k_0^2\right] 
+ \tfrac{1}{2}\KT (\Curl {\bf n})^2 + \KG  \det \nabla {\bf n} + \tfrac{1}{2}\KTH |{\bf t}|^2 \,dA.
\end{aligned}
\end{equation}
Here, 
%$(\nabla {\bf n})_{ij} = \partial_{x_j}{\bf n} \cdot {\bf e}_i$ is the surface gradient with
%orthonormal tangent vectors $\{{\bf e}_1, {\bf e}_2\}$.
%%\todo[inline]{YNY and BQ: What do you call a coordinate system where the metric is the identity at one point?}
%Also, the surface divergence 
%$\Div {\bf n}$ is the trace, and the surface curl $\Curl {\bf n}$ is the anti-trace (sum of the off-diagonal elements), of $\nabla {\bf n}$ respectively.
%respectively (Figure \ref{fig:distortions}A--C).
the deformations come with elastic coefficients: the bending modulus $\KB$, twist modulus $\KT$ 
and saddle-splay modulus $\KG$ and tilt modulus $\KTH$
%HK define the tilt vector ${\bf t} = {\bf n}/({\bf N}\cdot {\bf n}) - {\bf N}$ where ${\bf N}$ is the unit surface normal.
%
The parameter $k_0$ is the  spontaneous curvature and it determines the preferred lipid splay \cite{RoLi15,Kozlov2007}. 
As an illustration, the lipid DSPC has $k_0 = -0.1$ nm$^{-1}$ 
%as a result of  %having a relatively long acyl chain. 
and these lipids would line the  inner monolayer of a 
spherical liposome, because the addition of a positive splay $\Div \mathbf{n}$  in \eqref{ansatz3}
to the negative spontaneous curvature $k_0$ leads to lower energy \cite{Kamal22245, C3SM51829A, RoLi15,FriedSeguin15}.
The subtraction of $k_0^2$ makes the energy distortion free
\cite{Helfrich73,PhysRevLett.113.248102,Hamm2000}.

@@

%\begin{center}
%Anticipated model simulations
%\end{center}
\subsubsection{Simulations of the HARP model to estimate elastic moduli}
%To characterize material properties of a bilayer, we compare the energies and forces generated by
%a collection of amphiphilic particles against the energies and forces of a continuum bilayer.
We propose to
%
%to examine conduct calculations similar to the example of figure~\ref{fig:flattening}
%for a wide range of particle parameters and physical parameters in HARP to 
%
%{\bf (i)} Assess whether the continuum energies \eqref{ansatz3} are the same
%as the corresponding energies in the particle system for all reasonable deformation.
%If this is the case we will 
%
{\bf (i)} determine the value of the elastic moduli $\KB$,
$\KT$, $\KG$ and $\KTH$, and then {\bf (ii)} understand how the model parameters $\rho,$ $\gamma$, and particle shape map onto elastic moduli. 
An accurate and robust way to measure material properties is to track the evolution of the bilayer as
%
\begin{wrapfigure}[10]{r}{0.4\textwidth}
\vspace{-0.3cm}
\centerline{\includegraphics[width=0.4\textwidth]{Figures/Flattening.pdf}}
%  \centering{\includegraphics[width=0.80\textwidth]{Figures/Deformations.pdf}}
\vspace{-0.2cm}
\caption{\label{fig:flattening}Example of computing bending modulus of a lipid bilayer from particle simulation.
    }
\end{wrapfigure}
it relaxes from an initially non-equilibrium configuration.
Figure~\ref{fig:flattening} is an example where the initial configuration of a lipid bilayer patch contains only splay deformation without any other components (saddle splay, twist and tilt). 
As the bent membrane flattens we find that both the membrane self-interaction energy $\Phi$ and the elastic energy ${\cal W}$ decrease, and the slope gives the
splay modulus in this case.

>>>>

The HK framework assumes a three-dimensional lipid monolayers with internal structure
consisting of straight fibers that represent elongated hydrocarbon chains. HK assumed an
elastic energy density ${\cal W}$ that quadratic in the Green-Lagrange strain tensor for membrane deformation of
this striated, internal structure.
%
%
%
%These fibers
%tilt and stretch with respect to the dividing surface--the surface formed lying between the hydrocarbon chains and polar heads of lipids.
%Accordingly, the deformation takes the form 
%\begin{equation}
%  \label{LMdeformation}
%{\bf x}(X_1, X_2, X_3) = {\bf x}_0(X_1, X_2) + \zeta(x_0(X_1, X_2), X_3) {\bf n}(x_0(X_1, X_2))
%\end{equation}
%where $(X_1,X_2,X_3)$ are points in a reference volume.
%The map $x_0$ parametrizes the dividing surface $\Sigma$, and the unit vector field ${\bf n}$ parametrizes the
%direction of the lipid tails in the deformed state \cite{doi:10.1021/jp075641w,KLAUDA20083074}.
%The parameter $X_3$ and the function $\zeta$ parametrizes the distance along the hydrocarbon chain in the reference state
%and the deformed state, respectively, and so $0 \leq X_3 \leq \delta$ where $\delta = $ 2.5 nm is a realistic monolayer thickness. 
%
%The elastic theory assumes an energy density that is quadratic in the Green-Lagrange strain tensor for ${\bf x}$. 
%The chains stretch and compress to satisfy incompressibility.
%Symmetries with respect to mirror reflection and in-plane rotation, the internal structure assumptions and incompressibility
%lead to an elastic surface energy over the dividing surface $\Sigma$. 
%
This energy decomposes into four, fundamental and independent deformations  (figure~\ref{fig:deformations}):
splay ($\Div {\bf n}$), twist ($\Curl {\bf n}$), saddle splay ($\det \nabla {\bf n}$) and tilt ${\bf t}={\bf n}/({\bf N}\cdot {\bf n}) - {\bf N}$ where ${\bf N}$ is the unit surface normal:
\begin{equation}
\label{ansatz3}
\begin{aligned}
{\cal W} \equiv \int_{\Sigma} 
  \tfrac{1}{2}\KB\left[ \left( \Div {\bf n} + k_0\right)^2 - k_0^2\right] 
+ \tfrac{1}{2}\KT (\Curl {\bf n})^2 + \KG  \det \nabla {\bf n} + \tfrac{1}{2}\KTH |{\bf t}|^2 \,dA.
\end{aligned}
\end{equation}
Here, 
%$(\nabla {\bf n})_{ij} = \partial_{x_j}{\bf n} \cdot {\bf e}_i$ is the surface gradient with
%orthonormal tangent vectors $\{{\bf e}_1, {\bf e}_2\}$.
%%\todo[inline]{YNY and BQ: What do you call a coordinate system where the metric is the identity at one point?}
%Also, the surface divergence 
%$\Div {\bf n}$ is the trace, and the surface curl $\Curl {\bf n}$ is the anti-trace (sum of the off-diagonal elements), of $\nabla {\bf n}$ respectively.
%respectively (Figure \ref{fig:distortions}A--C).
the deformations come with elastic coefficients: the bending modulus $\KB$, twist modulus $\KT$ 
and saddle-splay modulus $\KG$ and tilt modulus $\KTH$
%HK define the tilt vector ${\bf t} = {\bf n}/({\bf N}\cdot {\bf n}) - {\bf N}$ where ${\bf N}$ is the unit surface normal.
%
The parameter $k_0$ is the  spontaneous curvature and it determines the preferred lipid splay \cite{RoLi15,Kozlov2007}. 
As an illustration, the lipid DSPC has $k_0 = -0.1$ nm$^{-1}$ 
%as a result of  %having a relatively long acyl chain. 
and these lipids would line the  inner monolayer of a 
spherical liposome, because the addition of a positive splay $\Div \mathbf{n}$  in \eqref{ansatz3}
to the negative spontaneous curvature $k_0$ leads to lower energy \cite{Kamal22245, C3SM51829A, RoLi15,FriedSeguin15}.
The subtraction of $k_0^2$ makes the energy distortion free
\cite{Helfrich73,PhysRevLett.113.248102,Hamm2000}.

@@

%\begin{center}
%Anticipated model simulations
%\end{center}
\subsubsection{Simulations of the HARP model to estimate elastic moduli}
%To characterize material properties of a bilayer, we compare the energies and forces generated by
%a collection of amphiphilic particles against the energies and forces of a continuum bilayer.
We propose to
%
%to examine conduct calculations similar to the example of figure~\ref{fig:flattening}
%for a wide range of particle parameters and physical parameters in HARP to 
%
%{\bf (i)} Assess whether the continuum energies \eqref{ansatz3} are the same
%as the corresponding energies in the particle system for all reasonable deformation.
%If this is the case we will 
%
{\bf (i)} determine the value of the elastic moduli $\KB$,
$\KT$, $\KG$ and $\KTH$, and then {\bf (ii)} understand how the model parameters $\rho,$ $\gamma$, and particle shape map onto elastic moduli. 
An accurate and robust way to measure material properties is to track the evolution of the bilayer as
%
\begin{wrapfigure}[10]{r}{0.4\textwidth}
\vspace{-0.3cm}
\centerline{\includegraphics[width=0.4\textwidth]{Figures/Flattening.pdf}}
%  \centering{\includegraphics[width=0.80\textwidth]{Figures/Deformations.pdf}}
\vspace{-0.2cm}
\caption{\label{fig:flattening}Example of computing bending modulus of a lipid bilayer from particle simulation.
    }
\end{wrapfigure}
it relaxes from an initially non-equilibrium configuration \cite{PhysRevLett.117.188102}.
Figure~\ref{fig:flattening} shows an example where the initial configuration of a lipid bilayer patch contains only
splay deformation without any other components (saddle splay, twist and tilt). 
As the bent membrane flattens we find that both the membrane self-interaction energy $\Phi$ and the elastic energy ${\cal W}$ decrease,
and because the saddle splay, twist and tilt stay zeros, the slope gives the bending modulus in this case.

### 2020-11-12 NSF2020/Description/specificaim2_ver2.tex
RJR

We include a background flow by replacing the third equation in \eqref{eq:stokes} with the condition
$(\mathbf{u} - \mathbf{u}_{\infty})(\mathbf{x}) \to \mathbf{0}$ as $|\mathbf{x}| \to \infty.$ 
One way to ensure this far-field condition is through the represention 
\begin{equation}
\label{PowerMiranda}
{\bf u} = {\bf u}_{\infty} + D\eta + \sum_{i=1}^N S(\cdot, {\bf a}_i) {\bf F}_i + R(\cdot, {\bf a}_i) {\bf G}_i.
\end{equation}
where $S$ and $R$ are stokeslets and rotlets supported at the respective particle centers 
and $D\eta$ is an appropriate double layer integral operator for an unknown density function $\eta.$ 
The representation \eqref{PowerMiranda} automatically satisfies all equations with the exception of the rigid motion boundary conditions. 
These conditions follow by requiring the viscous stress vanishes across the particle boundaries.

@@

\begin{wrapfigure}[16]{r}{0.32\textwidth}
\centerline{\includegraphics[width=0.32\textwidth]{figures/BIError.pdf}}
\caption{
\label{fig:bierror}  
  Gradient error between an exact and boundary integral solution of \eqref{SL}.
}
\end{wrapfigure}

@@

\textbf{Novel reciprocal relations.}
We use a double layer The velocity field  double layer potential representation.
This representation involves a strongly singular integral, and without the aid of specilized
quadrature routines, leads to large numerical errors when evaluating the the hydrophobic stress \eqref{stress} along the particle boundaries (see Figure \cite{fig:bierror}).
We prove a, to our knowledge new, result concerning the evaluation of Maxwell-type stresses. 
\begin{proposition}
  \label{prop:recip}
  Suppose that $u = \sum_{i=1}^N u_i$ where $u_i$ is a solution of the screened Laplace equation
  in $\mathbb{R}^3 \setminus P_i$ for $i=1,\dots, N.$ Then 
  \begin{equation}
    \label{eq:reciprocal}
{\bf F}_i = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_i = \sum_{j \neq i} \int_{\partial P_i} {\bf x} \times [\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, \quad i = 1\dots, N.
\end{equation}
where $\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i \cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump in stress across $\partial \Omega.$ 
\end{proposition}

>>>>

We include a background flow by replacing the third equation in \eqref{eq:stokes} with the condition
$(\mathbf{u} - \mathbf{u}_{\infty})(\mathbf{x}) \to \mathbf{0}$ as $|\mathbf{x}| \to \infty.$ 
The main idea to incorporate the far-field flow is to use the representation 
\begin{equation}
\label{PowerMiranda}
{\bf u} = {\bf u}_{\infty} + K\eta + \sum_{i=1}^N S(\cdot, {\bf a}_i) {\bf F}_i + R(\cdot, {\bf a}_i) {\bf G}_i.
\end{equation}
where $S$ and $R$ are stokeslets and rotlets supported at the respective particle centers 
and $K\eta$ is a layer potential for an unknown density function $\eta.$ 
The representation \eqref{PowerMiranda} automatically satisfies all equations, with the exception of the rigid motion conditions. 
These rigid motion conditions follow by requiring the viscous stress vanishes across the particle boundaries.

@@

\begin{wrapfigure}[13]{r}{0.32\textwidth}
\centerline{\includegraphics[width=0.32\textwidth]{figures/BIError.pdf}}
\caption{
\label{fig:bierror}  
The numerical error in integral representations blows up near boundaries.
}
\end{wrapfigure}

@@

\textbf{Novel reciprocal relations.}
We use a boundary integral representation for the velocity field. Due to singular behavior of the
kernels, the boundary integrals can have large numerical epresentation without the aid of specilized
quadrature routines, leads to large numerical errors when evaluating the the hydrophobic
stress \eqref{stress} along the particle boundaries (see Figure \cite{fig:bierror}).
We prove a, to our knowledge new, result concerning the evaluation of Maxwell-type stresses. 
\begin{proposition}
  \label{prop:recip}
  Suppose that $u = \sum_{i=1}^N u_i$ where $u_i$ is a solution of the screened Laplace equation
  in $\mathbb{R}^3 \setminus P_i$ for $i=1,\dots, N.$ Then 
  \begin{equation}
    \label{eq:reciprocal}
{\bf F}_i = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_i = \sum_{j \neq i} \int_{\partial P_i} {\bf x} \times [\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, \quad i = 1\dots, N.
\end{equation}
where $\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i \cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump in stress across $\partial \Omega.$ 
\end{proposition}
